---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
 
```{r}
Activity<-read.csv("activity.csv")
head(Activity) 
```

## What is mean total number of steps taken per day?
###Calculate total steps taken per day.
```{r}
TotalDailySteps<-aggregate(steps~date, Activity, sum)

```
###Place data into Histogram

```{r}
hist(TotalDailySteps$steps, xlab="Total Steps Per Day", ylab="Number of Days", main="Total Daily Steps")

```
###Calculate and report the mean and median of the total number of steps taken per day.
###Calculate the mean
```{r}
meanDailySteps<-mean(TotalDailySteps$steps)
meanDailySteps
```
###Calculate the median
```{r}
medianDailySteps<-median(TotalDailySteps$steps)
medianDailySteps
```

## What is the average daily activity pattern?
###Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
###Take average number of steps.
```{r}
AverageNumberSteps<-aggregate(steps~interval, Activity, mean)
```
###Make time series plot.
```{r}
with(AverageNumberSteps, plot(interval, steps, type = "l"))
```
###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
AverageNumberSteps[which.max(AverageNumberSteps[,2]),1]
```



## Imputing missing values
###Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
MissingValues<-is.na(Activity[,1])
```

```{r}
summary(MissingValues)
```
###2304 missing values

###Devise a strategy for filling in all of the missing values in the dataset. 

```{r}
AddedMissingValues <- AverageNumberSteps$Mean[match(Activity$interval, AverageNumberSteps$Interval)]
```
###Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
ActivityDataFilled<- transform(Activity,
                          steps = ifelse(is.na(Activity$steps), 
                yes = MissingValues, no = Activity$steps))

```
###new dataset with values added
```{r}
TotalFilledData <- aggregate(steps ~ date, ActivityDataFilled, sum)
```
```{r}
names(TotalFilledData) <- c("date", "dailySteps")

```


```{r}
sum(is.na(TotalFilledData$dailySteps))
```
###checks out.

###build data frame
```{r}
DFTotalFilledData <- data.frame(TotalFilledData)
```
###

```{r}
StepsDataTotal <- ggplot(DFTotalFilledData, aes(x = dailySteps)) + 
  geom_histogram(breaks = seq(0, 25000, by = 2500), fill = "green", col = "grey") + 
  ylim(0, 30) + 
  xlab("Total Steps Per Day") + 
  ylab("Frequency") + 
  ggtitle("Total Number of Steps Taken Daily")
  

print(StepsDataTotal)
```
###find mean
```{r}
mean(TotalFilledData$dailySteps)

```
###find median
```{r}
median(TotalFilledData$dailySteps)

```


## Are there differences in activity patterns between weekdays and weekends?
###make sure date is in appropriate format
```{r}
Activity$date <- as.Date(strptime(Activity$date, format="%Y-%m-%d"))
```

```{r}
Activity$dayType <- sapply(Activity$date, function(x) {
  if(weekdays(x) == "Saturday" | weekdays(x) == "Sunday")
  {y <- "Weekend"}
  else {y <- "Weekday"}
  y
})
```

```{r}
ActivityDay <-  aggregate(steps ~ interval + dayType, Activity, mean, na.rm = TRUE)

```
```{r}
PerDay <-  ggplot(ActivityDay, aes(x = interval , y = steps, color = dayType)) + 
  geom_line() + ggtitle("Average Steps Taken; Weekend or Weekday") + 
  xlab("Interval") + 
  ylab("Average Number of Steps") +
  facet_wrap(~dayType, ncol = 2, nrow=1) 
 
  

print(PerDay) 
```

